Level-Set method on \texttt{p4est}

### 2015-06-23 paper/tex/applications.tex
Carsten's comments addressed.

In this section we apply our approach to the study of the phase transition of a liquid melt to a solid crystaline structure. In the case of a single component melt, and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem. We decompose the computational domain $\Omega$ into two subdomains $\Omega_l$ and $\Omega_s$, separated by an interface $\Gamma$. The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega_s$ and $T_l$ in the liquid phase $\Omega_l$, as
\begin{align}
\pd{T_l}{t} & = D_l \Delta T_l \quad \mathrm{in} ~~ \Omega_l, \label{eq:stefan_heat_equation_l} \\
\pd{T_s}{t} & = D_s \Delta T_s \quad \mathrm{in} ~~ \Omega_s. \label{eq:stefan_heat_equation_s}
\end{align}
The diffusion constants $D_l$ and $D_s$ can be discontinuous across the interface. We prescribe homogeneous Neumann boundary conditions on the edge of the computational domain, $\nabla T \cdot \underline{\mathbf{n}}\vert_{\partial \Omega}=0$. At the interface between the solid and the liquid phases, the temperature is given by the Gibbs-Tompson boundary condition \cite{Alexiades;Solomon;Wilson:88:The-formation-of-a-s, Alexiades;Solomon:93:Mathematical-Modelin}:
\begin{equation} \label{eq:stefan_gibbs_tompson}
T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\underline{\mathbf{u}} \cdot \underline{\mathbf{n}}),
\end{equation}
where $\kappa$ is the local interface curvature, $\underline{\mbf{u}}$ is the velocity of the interface, $\underline{\mathbf{n}}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the surface tension and kinetic undercooling coefficients. The interface velocity $\underline{\mbf{u}}$ is defined from the jump in the heat flux at the interface,
\begin{equation} \label{eq:stefan_velocity}
(\underline{\mathbf{u}} \cdot \underline{\mathbf{n}}) = - \left[ D_l \pd{T_l}{\underline{\mathbf{n}}} - D_s \pd{T_s}{\underline{\mathbf{n}}} \right].
\end{equation}
We choose to use an adaptive time step with a $\text{CFL} = 5$, i.e.
\begin{equation} \label{eq:stefan_dt}
\Delta t = 5 ~ \Delta x_{min} ~ \min(1,1/\max \lVert \underline{\mathbf{u}} \rVert),
\end{equation}
where $\Delta x_{min}$ is the size of the smallest cell of the forest. The general procedure to solve the Stefan problem is presented in Algorithm \ref{alg:stefan}.

@@

The implementation of the Stefan problem relies on the components described in the previous sections, and it is therefore a good synthesis of the performance of the various algorithms. We monitored the performance of the code over five time iterations, as presented in Algorithm \ref{alg:stefan}, for two different maximum resolutions. In both cases, the forest is built on a $20\times20\times20$ macro-mesh. The maximum tree resolution for the small test is 9, leading to approximately 7M grid points, and the maximum resolution for the large test is 11, corresponding to 105M grid points. The results are presented in figure \ref{fig:stefan_scaling}. As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (cf.\ Table \ref{tab:scaling_stefan}).

>>>>

In this section we apply our approach to the study of the phase transition of a liquid melt to a solid crystaline structure. In the case of a single component melt, and in the absence of convection, the process is dominated by diffusion and can be modeled as a Stefan problem. We decompose the computational domain $\Omega$ into two subdomains $\Omega_l$ and $\Omega_s$, separated by an interface $\Gamma$. The Stefan problem describes the evolution of the temperature $T$, decomposed into $T_s$ in the solid phase $\Omega_s$ and $T_l$ in the liquid phase $\Omega_l$, as
\begin{align}
\pd{T_l}{t} & = D_l \Delta T_l \quad \mathrm{in} ~~ \Omega_l, \label{eq:stefan_heat_equation_l} \\
\pd{T_s}{t} & = D_s \Delta T_s \quad \mathrm{in} ~~ \Omega_s. \label{eq:stefan_heat_equation_s}
\end{align}
The diffusion constants $D_l$ and $D_s$ can be discontinuous across the interface. We prescribe homogeneous Neumann boundary conditions on the edge of the computational domain, $\nabla T \cdot \underline{\mathbf{n}}\vert_{\partial \Omega}=0$. At the interface between the solid and the liquid phases, the temperature is given by the Gibbs-Tompson boundary condition \cite{Alexiades;Solomon;Wilson:88:The-formation-of-a-s, Alexiades;Solomon:93:Mathematical-Modelin}:
\begin{equation} \label{eq:stefan_gibbs_tompson}
T_s = T_l = T_{\Gamma} = -\epsilon_c \kappa - \epsilon_v (\underline{\mathbf{u}} \cdot \underline{\mathbf{n}}),
\end{equation}
where $\kappa$ is the local interface curvature, $\underline{\mbf{u}}$ is the velocity of the interface, $\underline{\mathbf{n}}$ is the outward normal to the solidification front and $\epsilon_c$ and $\epsilon_v$ are the surface tension and kinetic undercooling coefficients. The interface velocity $\underline{\mbf{u}}$ is defined from the jump in the heat flux at the interface,
\begin{equation} \label{eq:stefan_velocity}
(\underline{\mathbf{u}} \cdot \underline{\mathbf{n}}) = - \left[ D_l \pd{T_l}{\underline{\mathbf{n}}} - D_s \pd{T_s}{\underline{\mathbf{n}}} \right].
\end{equation}
We choose to use an adaptive time step with a $\text{CFL} = 5$, i.e.
\begin{equation} \label{eq:stefan_dt}
\Delta t = 5 ~ \Delta x_{min} ~ \min(1,1/\max \lVert \underline{\mathbf{u}} \rVert),
\end{equation}
where $\Delta x_{min}$ is the size of the smallest cell of the forest. The general procedure to solve the Stefan problem is presented in Algorithm \ref{alg:stefan} and we refer the interested reader to \cite{Chen;Min;Gibou:09:A-numerical-scheme-f} for the details of implementation. In implementing the numerical solver, we make use of the popular \texttt{PETSc}
\cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.

@@

The implementation of the Stefan problem relies on the components described in the previous sections, and it is therefore a good synthesis of the performance of the various algorithms. We monitored the performance of the code over five time iterations, as presented in Algorithm \ref{alg:stefan}, for two different maximum resolutions. In both cases, the forest is built on a $20\times20\times20$ macro-mesh. The maximum tree resolution for the small test is 9, leading to approximately 7M grid points, and the maximum resolution for the large test is 11, corresponding to 105M grid points. The results are presented in figure \ref{fig:stefan_scaling}, where ``\verb|Solution_Extension|'' refers to extrapolation procedure (see Algorithm \ref{alg:stefan} step 5). As expected from the results obtained for each component in the previous sections, our implementation of the Stefan problem exhibits very satisfactory scaling (cf.\ Table \ref{tab:scaling_stefan}).

### 2015-06-23 paper/tex/parallel_algorithms.tex
Exemplarily use mathit in some places
Carsten's comments addressed.

Finally, in \texttt{p4est} trees are linearized, i.e.\ only the leaves are explicitly stored. However, explicit knowledge of the hierarchal structure of the tree is greatly beneficial in several algorithms, e.g. in search operations needed for the interpolation algorithm. Thus, we introduce a simple reconstruction algorithm that recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer. This approach is similar to the ideas introduced in \cite{Bangerth;Burstedde;Heister;etal:11:Algorithms-and-data-} and our tests show that in a typical application they amount to less that $1\%$ of the entire runtime. Algorithm \ref{alg:reconstruction} illustrates how this reconstruction is performed. Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of the forest by recursively refining from the root until reaching the same level and location of all leaves in the local forest and the ghost layer. Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced provided that the initial forest is balanced. Figure \ref{fig:reconstruction} illustrates an example where Algorithm \ref{alg:reconstruction} is applied. Note how each process has independently generated a local representation of the forest that is refined to match the same leaves as in the global forest and ghost layer.
\begin{figure}[htbp]
\begin{center}
\includegraphics[width = \textwidth]{figures/reconstruct.pdf}
\end{center}
\caption{Left: a forest refined close to an interface and partitioned among four processes, as indicated by colors. Right: each process independently recreates a local forest that is refined to match the local grid and is as coarse as possible elsewhere. Note that empty cells are fictitious, i.e.\ they are only required to generate the hierarchal structure and are not matched by any corresponding cell in the global forest.}
\label{fig:reconstruction}
\end{figure}
\begin{algorithm}[htbp]
\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$: construction of the local tree hierarchy $H$ from the grid $G$}
\begin{algorithmic}[1]
\State $H \gets G.\texttt{macromesh()}$
\For {$tr : G.\texttt{local\_trees()}$} \Comment{build hierarchy for local cells}
	\For {$c : tr.\texttt{cells()}$}
		\State $H.\texttt{update\_tree(} tr, c \texttt{)}$
	\EndFor
\EndFor

@@

\State \Return $H$
\\
\Function{$H.\texttt{update\_tree}$}{$tr, c$}   \Comment{recursive tree reconstruction}
	\State $c_l \gets H.\texttt{root(}tr\texttt{)}$
	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $}
		\If {$c_l.$\texttt{is\_leaf()}} $c_l$\texttt{.split()}
		\EndIf
		\State $h \gets c_l.\texttt{length()} / 2$ \Comment{select the next child based on direction}
		\State $i \gets c.x \ge c_l.x + h$
		\State $j \gets c.y \ge c_l.y + h$
		\State $k \gets c.z \ge c_l.z + h$

@@

To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e.\ we construct the matrix $A_{P \times P}$, where $P$ is the number of processes, such that
\ben
a_{ij} = 
\left\{
\begin{array}{lr}
1 \hspace{5 mm} \text{if process `\textit{i}' sends a message to process `\textit{j}',} \\
0 \hspace{5 mm} \text{otherwise.}
\end{array}
\right.
\een
Note that this matrix is also distributed among processes, i.e.\ each row is owned by a separate rank. Next, we compute
\ben
S_i = \sum_j a_{ij} \hspace{5 mm} \text {and} \hspace{5 mm} R_i = \sum_j a_{ji} ,
\een
where $S_i$ and $R_i$ denote the number of messages sent and received, respectively. While $S_i$ can be computed trivially, a reduction operation is required to compute $R_i$. For instance, this can be achieved using a single \texttt{MPI\_Reduce\_scatter} function call. The last phase of the interpolation procedure involves overlapping the computation of interpolated values for local points with the communication of data between processes. This is done by alternating between local calculations and probing for incoming messages from other processes. The interpolation is finished once the values for all local the points have been calculated and all the remote requests have been processed (see Algorithm \ref{alg:interpolation}).
\begin{algorithm}[htbp]
\caption{$values \gets \texttt{Interpolate (}H, F, \underline{\mathbf{X}}\texttt{)}$: interpolate the value of $F$, defined on the local tree hierarchy $H$, at coordinates $\underline{\mathbf{X}}$}
\begin{algorithmic}[1]
\State $col \gets 0, buff \gets null$ \Comment{Phase I -- buffering}
\For {$\underline{\mathbf{p}} : \underline{\mathbf{X}}$} 
	\State $[r, cell] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
	\If {$r = mpirank$}
		\State $buff[r].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$
	\Else
		\State $buff[r].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$
		\State $col[r] \gets 1$
	\EndIf
\EndFor
\For {$r: mpisize$} \Comment{Phase II -- initiate communication and compute number of messages}
	\If {$col[r]$}
		\State $\texttt{MP\_Isend(}r, buff[r]\texttt{)}$
	\EndIf
\EndFor
\State  $S \gets \texttt{sum(}col\texttt{)}$ 
\State  $R \gets \texttt{MPI\_Reduce\_scatter(}col,\texttt{MPI\_SUM)}$
\State $done \gets false, it \gets buff[mpirank].\texttt{begin()}$
\While {$!done$} \Comment{Phase III -- main loop}
	\If {$it \not= buff[mpirank].\texttt{end()}$}
		\State $values \gets \texttt{process\_local(}it\texttt{)}$ \Comment{process local interpolations}
		\State $\texttt{++}it$
	\EndIf
	\If {$R > 0$} \Comment{process queries sent from remote processes}
		\State $[msg, st] \gets \texttt{MPI\_Iprobe()}$
		\If {$msg$}
			% \State $values \gets \texttt{MPI\_Recv(}st.\texttt{MPI\_SOURCE)}$ \Comment{}
			\State $val\_buff \gets \texttt{process\_queries(}st\texttt{)}$ \Comment{receive, search, and interpolate values}
			\State $\texttt{MPI\_Isend(}st.\texttt{MPI\_SOURCE},val\_buff\texttt{)}$ \Comment{send back interpolated values}
			\State $R\texttt{--}$
		\EndIf
	\EndIf
	\If {$S > 0$} \Comment{process replies sent to our queries}
		\State $[msg, st] \gets \texttt{MPI\_Iprobe()}$
		\If {$msg$}
			\State $values \gets \texttt{process\_replies(}st.\texttt{MPI\_SOURCE)}$ \Comment{receive remotely interpolated values}
			\State $S\texttt{--}$
		\EndIf
	\EndIf
	\State $done \gets S = 0 \: \And \: R = 0 \: \And \: it = buff[mpirank].\texttt{end()}$
\EndWhile
\State \Return $values$
\end{algorithmic}
\label{alg:interpolation}
\end{algorithm}

Using the interpolation Algorithm \ref{alg:interpolation}, we close this section by presenting the final semi-Lagrangian Algorithm \ref{alg:semi-lagrangian}. The basic idea is to start from an initial guess $G^{n+1}_0$ for the grid and modify it using the refinement \eqref{eq:refine} and coarsening \eqref{eq:coarsen} criteria until convergence is obtained. Various options are available for $G^{n+1}_0$. For instance it is possible to start from the macromesh and only perform refinement steps until convergence. This choice, however, is not suitable since the first few iterations do not contain many cells and there is little work for parallelism. Here we simply take the previous grid as the starting point, i.e.\ $G^{n+1}_0 = G^n$. Note that this iterative process is essentially unavoidable since the grid is based on the values of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$. Nonetheless the process converges to the final grid in at most $l_{\max}-l_{\min}$ steps where $l_{\min}$ and $l_{\max}$ denote the maximum and minimum depth of all trees in the forest, receptively.
\begin{algorithm}[htbp]
\caption{$[G^{n+1}, \phi^{n+1}] \gets \texttt{SemiLagrangian (}G^n, \phi^n, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \text{CFL}\texttt{)}$: update $\phi^{n+1}$ from $\phi^n$ using a semi-Lagrangian scheme and construct the new forest $G^{n+1}$ that is consistent with the zero level-set of $\phi^{n+1}$}
\begin{algorithmic}[1]
\State $\Delta t_l \gets \text{CFL} \times G^n.\texttt{hmin()} / \max \{\underline{\mathbf{u}}^n \} $
\State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$
\State $H^n \gets \texttt{Reconstruct(} G^n\texttt{)}$ 
\State $G^{n+1}_0 \gets G^n$
\While {$true$}
	\State $\underline{\mathbf{X}}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \phi^n, \underline{\textbf{X}}_d\texttt{)}$
	\State $G^{n+1} \gets G^{n+1}_0.\texttt{refine\_and\_coarsen(}\phi^{n+1}\texttt{)}$ \Comment{using equations \ref{eq:refine} and \ref{eq:coarsen} as criteria}
	\If {$G^{n+1} \not= G^{n+1}_0$}
		\State $G^{n+1}.\texttt{partition()}$
		\State $G^{n+1}_0 \gets G^{n+1}$
	\Else
		\State \textbf{break}
	\EndIf
\EndWhile

@@

\begin{algorithm}[htbp]
\caption{$y \gets \texttt{Overlap (}x, \mathcal{F}\texttt{)}$: compute $y_i=\mathcal{F}(x_i)$ for all nodes $i$, where $\mathcal{F}$ is a local operation, while hiding the communication to update the ghost layer}
\begin{algorithmic}[1]
\For {$i:B_p$} \Comment{I -- perform computation on boundary points}
	\State $y_i \gets \mathcal{F}(x_i)$
\EndFor
\State $send\_req \gets \texttt{MPI\_Isend(}y_B\texttt{)}$ \Comment{II -- begin updating ghost values}
\State $recv\_req \gets \texttt{MPI\_Irecv(}y_G\texttt{)}$
\For {$i:P_p$} \Comment{III -- perform computation on private points}
	\State $y_i \gets \mathcal{F}(x_i)$ 
\EndFor
\State $\texttt{MPI\_Waitall(}send\_req, recv\_req\texttt{)}$ \Comment{IV -- wait for ghost update to finish}
\State \Return $y$
\end{algorithmic}
\label{alg:overlap}
\end{algorithm}

>>>>

Finally, in \texttt{p4est} trees are linearized, i.e.\ only the leaves are explicitly stored. However, explicit knowledge of the hierarchal structure of the tree is greatly beneficial in several algorithms, e.g. in search operations needed for the interpolation algorithm. Thus, we introduce a simple reconstruction algorithm that recreates a local representation of the entire ``forest'' that is only adapted to local cells and, potentially, the ghost layer. This approach is similar to the ideas introduced in \cite{Bangerth;Burstedde;Heister;etal:11:Algorithms-and-data-} and our tests show that in a typical application they amount to less that $1\%$ of the entire runtime. Algorithm \ref{alg:reconstruction} illustrates how this reconstruction is performed. Given a forest and a layer of ghost cells from \texttt{p4est}, the algorithm generates a local representation of the forest by recursively refining from the root until reaching the same level and location of all leaves in the local forest and the ghost layer. Note that algorithm \ref{alg:reconstruction} does not involve any communication and is load balanced provided that the initial forest is balanced. Figure \ref{fig:reconstruction} illustrates an example where Algorithm \ref{alg:reconstruction} is applied. Note how each process has independently generated a local representation of the forest that is refined to match the same leaves as in the global forest and ghost layer.
\begin{figure}[htbp]
\begin{center}
\includegraphics[width = \textwidth]{figures/reconstruct.pdf}
\end{center}
\caption{Left: a forest refined close to an interface and partitioned among four processes, as indicated by colors. Right: each process independently recreates a local forest that is refined to match the local grid and is as coarse as possible elsewhere. Note that empty cells are fictitious, i.e.\ they are only required to generate the hierarchal structure and are not matched by any corresponding cell in the global forest.}
\label{fig:reconstruction}
\end{figure}
\begin{algorithm}[htbp]
\caption{$H \gets \texttt{Reconstruct (}G\texttt{)}$: construction of the local tree hierarchy $H$ from the grid $G$}
\begin{algorithmic}[1]
\State $H \gets G.\texttt{macromesh()}$
\For {$\mathit{tr} : G.\texttt{local\_trees()}$} \Comment{build hierarchy for local cells}
	\For {$c : \mathit{tr}.\texttt{cells()}$}
		\State $H.\texttt{update\_tree(} \mathit{tr}, c \texttt{)}$
	\EndFor
\EndFor

@@

\State \Return $H$
\\
\Function{$H.\texttt{update\_tree}$}{$\mathit{tr}, c$}   \Comment{recursive tree reconstruction}
	\State $c_l \gets H.\texttt{root(}\mathit{tr}\texttt{)}$
	\While{$c_l.\texttt{level()} \not= c.\texttt{level()} $}
		\If {$c_l.$\texttt{is\_leaf()}} $c_l$\texttt{.split()}
		\EndIf
		\State $h \gets c_l.\texttt{length()} / 2$ \Comment{select the next child based on direction}
		\State $i \gets c.x \ge c_l.x + h$
		\State $j \gets c.y \ge c_l.y + h$
		\State $k \gets c.z \ge c_l.z + h$

@@

To solve the communication problem, we first compute the adjacency matrix of the communication pattern, i.e.\ we construct the matrix $A_{P \times P}$, where $P$ is the number of processes, such that
\ben
a_{ij} = 
\left\{
\begin{array}{lr}
1 \hspace{5 mm} \text{if process `\textit{i}' sends a message to process `\textit{j}',} \\
0 \hspace{5 mm} \text{otherwise.}
\end{array}
\right.
\een
Note that this matrix is also distributed among processes, i.e.\ each row is owned by a separate rank. Next, we compute
\ben
S_i = \sum_j a_{ij} \hspace{5 mm} \text {and} \hspace{5 mm} R_i = \sum_j a_{ji} ,
\een
where $S_i$ and $R_i$ denote the number of messages sent and received, respectively. While $S_i$ can be computed trivially, a reduction operation is required to compute $R_i$. For instance, this can be achieved using a single \texttt{MPI\_Reduce\_scatter} function call. The last phase of the interpolation procedure involves overlapping the computation of interpolated values for local points with the communication of data between processes. This is done by alternating between local calculations and probing for incoming messages from other processes. The interpolation is finished once the values for all local the points have been calculated and all the remote requests have been processed (see Algorithm \ref{alg:interpolation}).
\begin{algorithm}[htbp]
\caption{$values \gets \texttt{Interpolate (}H, F, \underline{\mathbf{X}}\texttt{)}$: interpolate the value of $F$, defined on the local tree hierarchy $H$, at coordinates $\underline{\mathbf{X}}$}
\begin{algorithmic}[1]
\State $col \gets 0, \mathit{buff} \gets \mathit{null}$ \Comment{Phase I -- buffering}
\For {$\underline{\mathbf{p}} : \underline{\mathbf{X}}$} 
	\State $[r, \mathit{cell})] \gets H.\texttt{search(}\underline{\mathbf{p}}\texttt{)}$ \Comment{search for the owner's rank and cell}
	\If {$r = \mathit{mpirank}$}
        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}, cell\texttt{)}$
	\Else
        \State $\mathit{buff}[r].\texttt{push\_back(}\underline{\mathbf{p}}\texttt{)}$
        \State $\mathit{col}[r] \gets 1$
	\EndIf
\EndFor
\For {$r:\mathit{mpisize}$} \Comment{Phase II -- initiate communication and compute number of messages}
	\If {$\mathit{col}[r]$}
        \State $\texttt{MP\_Isend(}r, \mathit{buff}[r]\texttt{)}$
	\EndIf
\EndFor
\State  $S \gets \texttt{sum(}col\texttt{)}$ 
\State  $R \gets \texttt{MPI\_Reduce\_scatter(}\mathit{col},\texttt{MPI\_SUM)}$
\State $\mathit{done} \gets \mathit{false}, it \gets \mathit{buff}[mpirank].\texttt{begin()}$
\While {$\mathit{!done}$} \Comment{Phase III -- main loop}
	\If {$\mathit{it} \not= \mathit{buff[mpirank]}.\texttt{end()}$}
		\State $\mathit{values} \gets \texttt{process\_local(}it\texttt{)}$ \Comment{process local interpolations}
		\State $\texttt{++}\mathit{it}$
	\EndIf
	\If {$R > 0$} \Comment{process queries sent from remote processes}
		\State $[\mathit{msg, st}] \gets \texttt{MPI\_Iprobe()}$
		\If {$\mathit{msg}$}
			% \State $values \gets \texttt{MPI\_Recv(}st.\texttt{MPI\_SOURCE)}$ \Comment{}
			\State $\mathit{val\_buff} \gets \texttt{process\_queries(}\mathit{st}\texttt{)}$ \Comment{receive, search, and interpolate values}
			\State $\texttt{MPI\_Isend(}\mathit{st}.\texttt{MPI\_SOURCE},\mathit{val\_buff}\texttt{)}$ \Comment{send back interpolated values}
			\State $R\texttt{--}$
		\EndIf
	\EndIf
	\If {$S > 0$} \Comment{process replies sent to our queries}
		\State $\mathit{[msg, st]} \gets \texttt{MPI\_Iprobe()}$
		\If {$\mathit{msg}$}
			\State $\mathit{values} \gets \texttt{process\_replies(}\mathit{st}.\texttt{MPI\_SOURCE)}$ \Comment{receive remotely interpolated values}
			\State $S\texttt{--}$
		\EndIf
	\EndIf
	\State $\mathit{done} \gets S = 0 \: \And \: R = 0 \: \And \: \mathit{it = buff[mpirank]}.\texttt{end()}$
\EndWhile
\State \Return $\mathit{values}$
\end{algorithmic}
\label{alg:interpolation}
\end{algorithm}

Using the interpolation Algorithm \ref{alg:interpolation}, we close this section by presenting the final semi-Lagrangian Algorithm \ref{alg:semi-lagrangian}. The basic idea is to start from an initial guess $G^{n+1}_0$ for the grid and modify it using the refinement \eqref{eq:refine} and coarsening \eqref{eq:coarsen} criteria until convergence is obtained. Various options are available for $G^{n+1}_0$. For instance it is possible to start from the macromesh and only perform refinement steps until convergence. This choice, however, is not suitable since the first few iterations do not contain many cells and there is little work for parallelism. Here we simply take the previous grid as the starting point, i.e.\ $G^{n+1}_0 = G^n$. Note that this iterative process is essentially unavoidable since the grid is based on the values of the level-set function at $t^{n+1}$, which itself is unknown and is to be defined on $G^{n+1}$. Nonetheless the process converges to the final grid in at most $l_{\max}-l_{\min}$ steps where $l_{\min}$ and $l_{\max}$ denote the maximum and minimum depth of all trees in the forest, receptively.
\begin{algorithm}[htbp]
\caption{$[G^{n+1}, \phi^{n+1}] \gets \texttt{SemiLagrangian (}G^n, \phi^n, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \text{CFL}\texttt{)}$: update $\phi^{n+1}$ from $\phi^n$ using a semi-Lagrangian scheme and construct the new forest $G^{n+1}$ that is consistent with the zero level-set of $\phi^{n+1}$}
\begin{algorithmic}[1]
\State $\Delta t_l \gets \text{CFL} \times G^n.\texttt{hmin()} / \max \{\underline{\mathbf{u}}^n \} $
\State $\Delta t   \gets \texttt{MPI\_Allreduce(}\Delta t_l,\texttt{MPI\_MIN)}$
\State $H^n \gets \texttt{Reconstruct(} G^n\texttt{)}$ 
\State $G^{n+1}_0 \gets G^n$
\While {$\mathit{true}$}
	\State $\underline{\mathbf{X}}_d  \gets \texttt{ComputeDeparturePoints(}G^{n+1}_0, \underline{\mathbf{u}}^n, \underline{\mathbf{u}}^{n-1}, \Delta t\texttt{)}$ \Comment{using equations \ref{eq:xstar} -- \ref{eq:vn_p_half}}
	\State $\phi^{n+1} \gets \texttt{Interpolate(}H^n, \phi^n, \underline{\textbf{X}}_d\texttt{)}$
	\State $G^{n+1} \gets G^{n+1}_0.\texttt{refine\_and\_coarsen(}\phi^{n+1}\texttt{)}$ \Comment{using equations \ref{eq:refine} and \ref{eq:coarsen} as criteria}
	\If {$G^{n+1} \not= G^{n+1}_0$}
		\State $G^{n+1}.\texttt{partition()}$
		\State $G^{n+1}_0 \gets G^{n+1}$
	\Else
		\State \textbf{break}
	\EndIf
\EndWhile

@@

\begin{algorithm}[htbp]
\caption{$y \gets \texttt{Overlap (}x, \mathcal{F}\texttt{)}$: compute $y_i=\mathcal{F}(x_i)$ for all nodes $i$, where $\mathcal{F}$ is a local operation, while hiding the communication to update the ghost layer}
\begin{algorithmic}[1]
\For {$i:B_p$} \Comment{I -- perform computation on boundary points}
	\State $y_i \gets \mathcal{F}(x_i)$
\EndFor
\State $\mathit{send\_req} \gets \texttt{MPI\_Isend(}y_B\texttt{)}$ \Comment{II -- begin updating ghost values}
\State $\mathit{recv\_req} \gets \texttt{MPI\_Irecv(}y_G\texttt{)}$
\For {$i:P_p$} \Comment{III -- perform computation on private points}
	\State $y_i \gets \mathcal{F}(x_i)$ 
\EndFor
\State $\texttt{MPI\_Waitall(}\mathit{send\_req, recv\_req}\texttt{)}$ \Comment{IV -- wait for ghost update to finish}
\State \Return $y$
\end{algorithmic}
\label{alg:overlap}
\end{algorithm}
